Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-01-15 manuscript/supplementaryRW.Rmd
Add modelling protocol for Formind

>>>>

Samuel M. Fischer, Friedrich Bohn, Rico Fischer

<!--
@article{fischer_lessons_2016,
	title = {Lessons learned from applying a forest gap model to understand ecosystem and carbon dynamics of complex tropical forests},
	volume = {326},
	issn = {03043800},
	url = {https://linkinghub.elsevier.com/retrieve/pii/S0304380015005505},
	doi = {10.1016/j.ecolmodel.2015.11.018},
	language = {en},
	urldate = {2022-09-15},
	journal = {Ecological Modelling},
	author = {Fischer, Rico and Bohn, Friedrich and Dantas de Paula, Mateus and Dislich, Claudia and Groeneveld, J{\"u}rgen and Guti{\'e}rrez, Alvaro G. and Kazmierczak, Martin and Knapp, Nikolai and Lehmann, Sebastian and Paulick, Sebastian and P{\"u}tz, Sandro and R{\"o}dig, Edna and Taubert, Franziska and K{\"o}hler, Peter and Huth, Andreas},
	month = apr,
	year = {2016},
	pages = {124--133},
	file = {Volltext:C\:\\Users\\fischsam\\Zotero\\storage\\EISNTZF6\\Fischer et al. - 2016 - Lessons learned from applying a forest gap model t.pdf:application/pdf},
}
-->

A full description of the model can be found in @fischer_lessons_2016. 
The basis for the parameterization we applied is described in @bohn2014.

To incorporate climate, we used the data set `daily_era5_chelsa_200.csv',
which contains climate data on a daily resolution. We considered the
covariates precipitation, temperature, and irradiance, the latter
of which we converted to units used in Formind. Adding on to these
data, we used a constant and site-independent potential evapotranspiration
(PET) value of $4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day
length was $12\text{h}$ at all sites.

Since we simulated longer time periods than the $38$ years for which
we had climate data, we used the provided data in a randomized fashion.
For each simulation year, we applied the climate data of a randomly
selected year. We neglected the existence of leap years, which led
to a minor time shift in climate data of later years.

We modelled soil conditions by using the provided soil quality values
as proxy for soil depth. To that end, we scaled the provided values
to the typical range of soil depth values used in Formind for temperate
forests. Specifically, we multiplied the soil quality values by factor 
$0.5$ to obtain the modelled soil depth in meters.

We did not use any topograpy data.

We considered each of the $11$ species in the protocol explicitly.
We chose the model parameters for the individual species based on
the work by @bohn2014. For species that were not
covered by @bohn2014, we used parameter values of
similar species with known parameters. The species composition was
not predefined and evolved solely from the seed input
and the model dynamics.

We simulated a forest area of $1\text{ha}$, which was divided into
$25$ patches of size $20\text{m}\times20\text{m}$. No interactions
between trees from different patches were considered in the model.
That is, the $25$ patches were independent replicates of one another.

#### Seed generation and dispersal

We modelled seed generation as global external seed rain. Specifically, we 
assumed that for each species, there is a constant seed inflow of 
$250\frac{\text{seeds}}{\text{ha}\cdot \text{yr}}$, which is distributed 
evenly among the $25$ individual patches. We supposed that all 
seeds promptly establish to seedlings with initial stem diameters at 
breast height (DBH) of 5\text{cm}.

#### Simulation length and sampling approach

We started the simulation at each site with a burn-in phase of $1000$
years and collected samples of the output variables in subsequent
years. The length of the burn-in phase was chosen based on earlier
experiences with the model. To confirm that the model reached its
limiting behaviour, we plotted the evolution of the output variables
(e.g. the basal area) at selected sites over time and verified that
no trends were visible after the burn-in phase.

We computed the output variables in equidistant time intervals of
$60$ years. Before taking a sample, we saved parts of the model's
state. Then, we simulated the forest for $10$ years and computed
the output variables, comparing the model's current state with the
state saved earlier (see below).
After taking the sample, we simulated the forest for another $50$
years to reduce correlations between samples. Then, we repeated the
procedure until the desired number of $200$ samples was collected.

To determine the number and basal area of recruits, we compared the
set of large trees before and after the regeneration time period of
$10$ years. Prior to simulating the forest for the regeneration time
period, we saved the internal IDs of all trees with DBHs above the
thresholds of $0.07\text{m}$ or $0.1\text{m}$, respectively. We
stored these IDs in set data structures $S_{0.07}^{\text{prior}}$ and
$S_{0.1}^{\text{prior}}$, corresponding to one threshold value each.
After the regeneration time, we repeated the procedure, obtaining
tree ID sets $S_{0.07}^{\text{posterior}}$ and $S_{0.1}^{\text{posterior}}$.
We then computed the set differences 
$S_{0.07}^{\text{recruits}}=S_{0.07}^{\text{posterior}}\backslash S_{0.07}^{\text{prior}}$
and $S_{0.1}^{\text{recruits}}=S_{0.1}^{\text{posterior}}\backslash S_{0.1}^{\text{prior}}$
and determined both the number and cumulative basal area of the trees
in $S_{0.07}^{\text{recruits}}$ and $S_{0.1}^{\text{recruits}}$ by considered
species, respectively.

Finally, we computed the total basal area for each considered species
by adding up the basal areas of all corresponding individuals. Note
that the initial DBH of seedlings in our model was $0.05\text{m}$.
Consequently, smaller trees were not included in the basal area computation.
